The watershed transform in ITK - discussion and new developments

### 2006-02-20 article/Article.tex
add Refnames to bibliography
darcs-hash:20060111191903-2fc9d-56da7d5590030781b82d5b922c238f7d42c7abb9.gz
fix title
darcs-hash:20060111204358-2fc9d-eb8ddc6da150ce0549440d0e37939efe1d58db78.gz
figure labels, basic color overlay generation, current itk watershed example, plateau image.
darcs-hash:20060118034932-fafb9-a957ed4bca9e80dd37374a0f69e98624d96866d9.gz
More details on hill climbing.
darcs-hash:20060220010825-fafb9-7ec8dff7f53cbcd544ae90efa7b241f696a76ef0.gz

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{subfigure}
\title{Finding regional extrema - methods\\ and performance}

@@

@@

An Obvious alternative that is very similar in practice is to 
smooth the image so that there are fewer regional minima.
\begin{figure}
\begin{center}
\label{fig:overseg}
\caption{Classical over segmentation.}
\end{center}
\end{figure}

@@

This can be achieved by performing a {\em reconstruction by erosion}
or {\em recursive conditional erosion} from the seleted regional
minima. Figure \ref{fig:recon} illustrates the results of apply a
reconstruction by erosion to a 1D image. The original image is shown
in blue and has regional minima A, B, C, D and E. The marker image
which is recursively eroded is shown in black, with a single zero at
regional minimum B. The red plateaus illustrate the parts of the
result image that are above the orginal. The resulting image has only
on regional minima that corresponds to B.
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{recon}
\label{fig:recon}
\caption{Reconstruction by erosion from a single regional minimum.}
\end{center}
\end{figure}

@@

Hill climbing requires a label image which is initialized to the
indexes of the regional minima at the regional minima and an UNKNOWN
value elsewhere\footnote{In the ITK implementation a separate image is
used to track UNKNOWN locations because UNKNOWN values are difficult
to deal with in a type independent way.}. The priority queue is
initialised with the borders of the regional minima. Pixels pushed
onto the queue with priority defined by their grey level and labelled
as they are pushed. When pixels are popped from the queue their
neighbors are checked, labelled, and pushed onto the queue if they
aren't watershed pixels. The lower completion step is carried out
implicitly by not allowing pixels to be pushed onto the queue with a
priority lower than the priority of the neighbor which was popped from
the queue.

The assumption that the distance between a pixel and all of its
neighbors is 1 means that the steepest upper neighbors of a pixel are
simply the unlabelled neighbors.

@@

\bibliographystyle{plain}
\bibliography{InsightJournal}
\nocite{ITKSoftwareGuide}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{subfigure}

\title{The watershed transform in ITK - discussion and new developments}

@@

The topological watershed appears to have been developed with the aim
of using measures of border significance for merging criteria. This
work lead to the discovery that it wasn't possible to compute certain
measures with the flooding approaches used by many implementations.

@@

An Obvious alternative that is very similar in practice is to 
smooth the image so that there are fewer regional minima.
\begin{figure}
\begin{center}
\caption{Classical over segmentation.}
\label{fig:overseg}
\end{center}
\end{figure}

@@

This can be achieved by performing a {\em reconstruction by erosion}
or {\em recursive conditional erosion} from the seleted regional
minima. Figure \ref{fig:recon} illustrates the results of apply a
reconstruction by erosion to a 1D image. The original image is shown
in blue and has regional minima A, B, C, D and E. The marker image
which is recursively eroded is shown in black, with a single zero at
regional minimum B. The red plateaus illustrate the parts of the
result image that are above the orginal. The resulting image has only
on regional minima that corresponds to B.
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.5]{recon}
\caption{Reconstruction by erosion from a single regional minimum.}
\label{fig:recon}
\end{center}
\end{figure}

@@

There seem to be at least 3 algorithms that fall into the
hill-climbing category.

%Hill climbing requires a label image which is initialized to the
%indexes of the regional minima at the regional minima and an UNKNOWN
%value elsewhere\footnote{In the ITK implementation a separate image is
%used to track UNKNOWN locations because UNKNOWN values are difficult
%to deal with in a type independent way.}. The priority queue is
%initialised with the borders of the regional minima. Pixels pushed
%onto the queue with priority defined by their grey level and labelled
%as they are pushed. When pixels are popped from the queue their
%neighbors are checked, labelled, and pushed onto the queue if they
%aren't watershed pixels. The lower completion step is carried out
%implicitly by not allowing pixels to be pushed onto the queue with a
%priority lower than the priority of the neighbor which was popped from
%the queue.

The assumption that the distance between a pixel and all of its
neighbors is 1 means that the steepest upper neighbors of a pixel are
simply the unlabelled neighbors. This applies to all 3 algorithms
mentioned below.

\subsubsection{Algorithm 1}
\begin{itemize}
\item lower complete the image
\item initialize the queue with all pixels except interior of minima.
\item while (!empty(queue))
   \begin{itemize}
	\item Take the pixel $p$ from the head of the queue
        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
        \begin{itemize}
         \item  if $q$ is unlabeled, then

propagate label$(p)$ to label$(q)$;

                else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then

                  label($q$) = WS;
	\end{itemize}

   \end{itemize}
\end{itemize}

There are at least two reasons why we shouldn't use this
approach. Firstly, the lower completion step is explicit, which we'd
like to avoid and secondly ALL non minima pixels are placed on the
queue, which means that the queue length is almost the same as the
number of image pixels, increasing the overhead of queue
management. We shall see that it is only necessary to have the pixels
on the borders of the regions in the queue.

In addition to these problems there is no provision for implicit
imposition of minima for a marker based segmentation approach.

\subsubsection{Algorithm 2}
Using ordered queues.
\begin{itemize}
\item initialize the queue with edges of the regional minima. 
The edges are pixels that are part of a regional minima with
neighbours that are background.
\item while (!empty(queue))
   \begin{itemize}
	\item Take the pixel $p$ from the head of the queue
        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
         \item  if $q$ is unlabeled then

     		propagate label$(p)$ to label$(q)$;

		place $q$ in the queue

                else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then

                label($q$) = WS;
	\end{itemize}

\end{itemize}

The queue only contains the borders of the growing regions.

\subsubsection{Algorithm 3}
Also using ordered queues.
\begin{itemize}
\item initialize the queue with background pixels that have 
      neighbors that are members of a regional minima. In practice
      this is done by checking all regional minima pixels to see
      whether they have background neighbors. The background pixels
      are placed on the queue and the label of the regional minima
      associated with them.
\item while (!empty(queue)
   \begin{itemize}
	\item Take the pixel $p$ from the head of the queue

         if $p$ is unlabelled;

           apply the associated label

         else

           check whether $p$ is a watershed pixel. $p$ is a watershed
           pixel if there is more than one label in the neighborhood.

        \item for all pixels $q$ that are steepest upper neighbors of $p$; do

	if $q$ is not in the queue;

           place in queue, associated with label($p$)

           mark as queued
   \end{itemize}
\end{itemize}

The difference between algorithms 2 and 3 is the order in which
labelling and queueing happens.

Algorithm 3 requires slightly more comlexity to store label
information in the queue.

{\em a summary of our findings comparing the last 2 algorithms}

@@

\bibliographystyle{plain}
\bibliography{InsightJournal,Refnames.bib}
\nocite{ITKSoftwareGuide}
